\chapter{Accretion Disks and Their Associated Outflows}

% \epigraph{``And now, I shall talk about the most efficient process in the universe. No, not bureaucracy..."}
% {--- \textup{Petri Savoleinen}, A talk at the Harvard-Smithsonian CfA}

As described in the Introduction, there are a wide variety of accreting systems
with varying degrees of astrophysical significance. Here I will describe the 
physics of accretion in more detail, before discussing the theoretical and observational 
basis for accretion disk winds. 



\section{Accretion Disks}

\subsection{The Physics of Accretion}

The basic phenomenon of accretion- matter falling into a gravitational potential well- 
is a ubiquitous one in astrophysics. The details of how and where the energy is released
and how angular momentum is transported is subject to a number of different 
interpretations, mainly depending on the {\em geometry} of the accretion flow.
The so-called $\alpha$-disk model developed by \cite{shakurasunyaev1973} is
currently the leading candidate for explaining how energy and angular momentum
is transported through a thin disk of material, an accretion disk.

By considering the energy released through viscous dissipation 
in the disk it is possible to derive a temperature distribution as a function of 
radius \citep{shakurasunyaev1973, fkrbook}. 

\begin{equation}
T(R) =  
\end{equation}

It is important to recognise that the work of \cite{shakurasunyaev1973} 
{\sl does not specify the nature of the disk SED}. What it does do is 
say where energy is originally released. Typically,
accretion disks are modelled as a series of annuli each emitting 
as blackbodies, but it is possible that a disk atmosphere with frequency-dependent
opacity would create a somewhat different spectrum. It is also possible that {\em neither} of these 
treatments are realistic. We shall therefore devote a little time to discussing
the observational arguments for accretion disks and the current problems 


\section{Observational Appearance}



\subsubsection{Potential Problems with the Thin-disk model}

A number of issues have been raised with the thin-disk model and
its applicability to accreting systems. 

\subsubsection{Quasar emission region sizes from microlensing}

\subsubsection{Quasar emission region sizes from X-ray lags}

\subsubsection{The 1000\AA break in AGN}



\subsubsection{The Spectral shape of CV disks}

Attempts to fit the observed SEDs of high-state CVs with simple disk models have met with mixed success. In
particular, the SEDs predicted by most stellar/disk atmosphere models 
are too blue in the UV \citep{wade1988,long1991,long1994,knigge1998} and exhibit
stronger-than-observed Balmer jumps in absorption 
\citep{wade1984,haug1987,ladous1989b,knigge1998}. One possible
explanation for these problems is that these models fail to capture
all of the relevant physics. Indeed, it has been argued that a
self-consistent treatment can produce better agreement with 
observational data (e.g. Shaviv et al. 1991;  but see also Idan et al. 2010).
\nocite{idanshaviv2010} \nocite{shaviv1991}
However, an alternative explanation, suggested by Knigge et al.
(1998b; see also Hassall et al. 1985)\nocite{KLWB98,hassall}, 
is that recombination continuum emission from the base of the 
disk wind might fill in the disk's Balmer absorption edge and flatten the UV spectrum.

\section{The Universality of Accretion}

Accretion appears to be an important physical processes across $\sim9$ orders
of magnitude in mass. But is this process the same at all scales? Does any 
behaviour manifest in all accretion systems? 

\subsection{The RMS-flux relation}


\subsection{Accretion States}


\subsection{Jets and Outflows}


\subsection{A Global Picture}

Clearly, accretion physics is relevant to a plethora of astrophysical phenomena. 
It would also appear that the outflowing material observed in accreting systems 
has a profound effect on the accretion process itself, as well as acting 
as a spectral `filter' -- modifying, and sometimes dominating the observational 
appearance of accretion disks.












%%%%WINDS 
%%% COULD MAKE THIS A NEW CHAPTER
\section{Accretion Disk Winds: Observational Evidence}

\section{Accretion Disk Winds: Driving Mechanisms}

\subsection{Thermal Winds}

\subsection{Line-driven Winds}

\subsection{Magneto-centrifugal Winds}

\section{Accretion Disk Wind Models}

\section{Wider Perspective}

